---
title: "A Short Introduction to Networks and Model Comparisons"
author: |
 | Levi Lee 
 | _
 | Amy Wagaman
 | Amherst College
date: November 16, 2016

output:
  beamer_presentation:
    theme: "Boadilla"
    colortheme: "orchid"
    fonttheme: "structurebold"
---

```{r, include=FALSE}
#output:
#  beamer_presentation:
#    theme: "Boadilla"
#    colortheme: "beaver"
#    fonttheme: "structurebold"

#output:
#  pdf_document:
#    fig_height: 3
#    fig_width: 10
#  html_document:
#    fig_height: 3
#    fig_width: 5
#  word_document:
#    fig_height: 3
#    fig_width: 5

library(gridExtra)
library(sand)
library(igraph)
library(combinat)
library(mdsr)
library(ggplot2)
library(xtable)

options(xtable.comment = FALSE)
```

# What is a Network? 

 - Using terminology from graph theory, a graph G = (V, E) is a structure that consists of a set V of vertices (nodes, actors, etc.), and a set E of edges (links, relationships, etc.)

 - Able to display multiple layers of data 

 - Four major groups: technological, biological, social, informational

# Example: Karate Club of Zachary (1977) 

```{r, echo=FALSE, fig.height=5, fig.align='center'}
data(karate)
plot(karate)
```

```{r, include=FALSE}
vcount(karate) #34 people
ecount(karate) #78 friednships

average.path.length(karate) #2.4082
diameter(karate) #13
transitivity(karate) #0.2556818 clustering coefficent
```

# Can we simulate this? 

 - What graph models/graph generators will we choose? 

 - What characterstics of the observed graph will we choose to look at?

 - What are some methods to access accuracy? 

# Erdos-Renyi Model

A graph model/generator with two parameters 

 - N: the number of vertices 

 - p: the probability of a link forming between any two pairs of nodes

```{r, echo=FALSE}
set.seed(490)

numsim <- 100 

g.er.karmod.coef <- rep(NA, numsim)
g.er.karmod.apl <- rep(NA, numsim)
g.er.karmod.dia <- rep(NA, numsim)

for (i in 1:numsim) {
  # n = number of vertices, p = probability of a link
  #p is caluclated as number of edges over number of possible edges
  #p is then (34 choose 2) since there is the possibility of a link between
  g.er.karmod <- erdos.renyi.game(n = vcount(karate), 
                                  p = (ecount(karate)/choose(vcount(karate), 2)))
  g.er.karmod.coef[i] <- transitivity(g.er.karmod)
  g.er.karmod.apl[i] <- average.path.length(g.er.karmod)
  g.er.karmod.dia[i] <- diameter(g.er.karmod)
}
```

```{r, echo=FALSE, fig.align='center', fig.height=4, message=FALSE}
set.seed(490)
g.er.karmod1 <- erdos.renyi.game(n = vcount(karate), 
                                 p = (ecount(karate)/choose(vcount(karate), 2)))

plot(g.er.karmod1)  
```

# Simulation

```{r, eval=FALSE, echo=FALSE, fig.align='center', fig.height=5, fig.width=5}
# lattice reference only; does not print 
par(mfrow=c(1,3))
hist(g.er.karmod.coef)
abline(v = transitivity(karate), col="red")


hist(g.er.karmod.apl)
abline(v = average.path.length(karate), col="red")

hist(g.er.karmod.dia)
abline(v = diameter(karate), col="red")
```


```{r, echo=FALSE, fig.align='center', fig.height=2, fig.width=5, warning=FALSE}
gplot1 <- ggplot(data = as.data.frame(g.er.karmod.coef), 
       aes(x = g.er.karmod.coef)) +
  geom_histogram(binwidth=0.0065, 
       colour="blue", 
       fill="lavender") + 
  geom_vline(xintercept = transitivity(karate), colour="red") + 
  labs(title="Simulated 
       Clustering Coefficients") + 
  xlab("Clustering Coefficient") + 
  ylab("Frequency") + 
  theme(legend.position="none", 
        plot.title = element_text(size=6),
        axis.text = element_text(size=4),
        axis.title = element_text(size=4)) 


gplot2 <- ggplot(data = as.data.frame(g.er.karmod.apl), aes(x = g.er.karmod.apl)) +
  geom_histogram(binwidth=0.031,
                 colour="blue",
                 fill = "lavender") + 
  geom_vline(xintercept = average.path.length(karate), colour="red") + 
  labs(title="Simulated 
       Average Path Lengths") + 
  xlab("Average Path Length") +
  ylab("") + 
  theme(legend.position="none", 
        plot.title = element_text(size=6),
        axis.text = element_text(size=4),
        axis.title = element_text(size=4)) 

gplot3 <- ggplot(data = as.data.frame(g.er.karmod.dia), aes(x = g.er.karmod.dia)) +
  geom_histogram(binwidth=0.12,
                 colour="blue",
                 fill = "lavender") + 
  geom_vline(xintercept = diameter(karate), colour="red") + 
  labs(title="Simulated Diameters") + 
  xlab("Diameter") + 
  ylab("") + 
  theme(legend.position="none", 
        plot.title = element_text(size=6),
        axis.text = element_text(size=4),
        axis.title = element_text(size=4))

grid.arrange(gplot1, gplot2, gplot3, ncol=3)
```

# What's Next?

Choose a different model!

 - Barabasi-Albert, Watts-Strogatz, ERGMs, R-MAT, HOT, COLD, ... 
 
Look at other properties of networks

 - Degree distributions, centrality measures, hub-like nodes, ...
 
# Sources

Chakrabarti, Deepayan, and Christos Faloutsos. "Graph mining: Laws, generators, and algorithms." ACM computing surveys (CSUR) 38.1 (2006): 2.

Kolaczyk, Eric D. Statistical Analysis of Network Data: Methods and Models. New York: Springer, 2009. Print.

Kolaczyk, Eric D., and Gabor Csardi. Statistical Analysis of Network Data with R. New York: Springer, 2014. Print.


