---
title: "A Short Introduction to Networks and Model Comparisons"
author: |
 | Levi Lee 
 | Advisor: Amy Wagaman
 | Amherst College
date: April 12, 2017

output:
  beamer_presentation:
    theme: "Warsaw"
    colortheme: "orchid"
    fonttheme: "professionalfonts"
---

```{r, include=FALSE}
library(sand)
library(igraph)
library(network)
library(sna)
library(statnet)
library(ergm)
library(xtable)
library(mdsr)
library(gridExtra)

options(xtable.comment = FALSE)
```

# A Highly Connected World 

 - Networks are everywhere 

# History 

 - Travers & Milgram (1967)
  
   - Letter correspondence between strangers in Nebraska and Massachusetts 
  
   - Overall, it took only around six people for the letter to be delivered 
  
   - "Six degrees of separation"

# Social Networks

 - Involves a lot of people 

 - Highly condensed groups

 - Relatively short distances between people 

# Question

 - Can we simulate this? If so, how? 
 
 - Conduct a simulation study utilizing different graph models 

# Basic Terminlogy

 - A *graph*, denoted $G(V, E)$, consists of a set of *vertices* $i, j, k, ... \in V$ and a set of *edges* $\{i, j\}, \{i, k\}, \{j, k\}... \in E$. 
 
 - Our focus is on *simple*, *undirected*, and *connected* graphs. 

# Other Representations of Graphs

 - *Adjacency matrix*: denoted $\textbf{A}$ is an $N_V$ x $N_V$ matrix where each element denotes the existence of edges between pairs where

$$
A_{ij} = \begin{cases}
    1 & \text{if } \{i, j\} \in E \\
    0 & \text{otherwise.} 
  \end{cases}
$$

 - *Edge list*: two-column list of all the edges in a graph denoted by their corresponding vertices present

# Example: Karate Club of Zachary (1977) 

```{r, echo=FALSE, fig.height=4.5, fig.align='center'}
data(karate)
plot(karate)
```

# Network Statistics 

 - Analogous to statisitcs seen in elementary statistics

 - Characterizes a given network

# Transitivity/Clustering Coefficient
 
 - Ratio of triangles to connected triples
   
 - *Triangle*: three vertices connected by three edges 
 
 - *Connected triple*: three vertices connected by two edges

 $$C = \frac {(\text{number of triangles}) \times 3} {\text{number of connected triples}}$$

# Notions of Distance 

*Average path length*: average of the shortest paths of all distinct pairs of vertices in the network

*Diameter*: longest of all the shortest paths between distinct pairs of vertices

# Example: Karate Club of Zachary and  Lazega's Law Firm
  -------------------------------------------------------
  Network Statistic               Zachary's Karate Club 
  ------------------------------- ---------------------- 
  Transitivity                    0.256            
    
  Average Path Length             2.408     
  
  Diameter                        13                    
  -------------------------------------------------------
  
# Centrality

 - Measure of importance for each vertex in the graph
 
 - Many different types of centralities exist 

# Degree Centrality 

 - Based on the number of edges are connected to a vertex 

 - Vertices with higher vertex degrees are considered to be more central to the network than those with lower vertex degrees

# Closeness Centrality 

 - Measures how close a vertex is to other vertices based on the inverse of the total distance of the vertex from all others
 
 $$c_{Cl}(i) = \frac {1} {\sum_{j \in V}^{} d(i, j)}$$
 
 - $dist(i,j)$ is the geodesic distance between the vertices $i,j \in V$

# Betweenness Centrality 

 - Measures the extent to which a vertex is located between other pairs of vertices
 
 $$c_{B}(i) = \sum_{g \neq h \neq i \in V}^{} \frac {\sigma(g,h|i)} {\sigma(g,h)}$$

 - $\sigma(g,h|i)$ is the total number of shortest paths between $g$ and $h$ that pass through $i$, and $\sigma(g,h) = \sum_{i \in V}^{} \sigma(g,h|i)$

# Eigenvector Centrality 

 - Based on the idea of "status," "prestige," or "rank;" the more central the neighbors of a vertex are, the more central that vertex itself is
 
 $$c_{Ei}(i) = \alpha \sum_{\{i,j\} \in E}^{} c_{Ei} (u)$$

 - $c_{Ei} = (c_{Ei}(1), ..., c_{Ei}(N_V))^T$ is the solution to the eigenvalue problem $\textbf{A}\textbf{c}_{Ei} = \alpha^{-1}\textbf{c}_{Ei}$, where $\textbf{A}$ is the adjacency matrix for network graph $G$.

# Example: Karate Club of Zachary (1977) 

```{r, echo = FALSE, fig.align = 'center', warning = FALSE}
A <- get.adjacency(karate, sparse=FALSE)
library(network)
g <- network::as.network.matrix(A)
library(sna)
par(mfrow=c(2,2))
sna::gplot.target(g, degree(g), main="Degree",
                  circ.lab = FALSE, circ.col="skyblue",
                  usearrows = FALSE,
                  vertex.col=c("blue", rep("red", 32), "yellow"),
                  edge.col="darkgray")


sna::gplot.target(g, closeness(g), main="Closeness",
                  circ.lab = FALSE, circ.col="skyblue",
                  usearrows = FALSE,
                  vertex.col=c("blue", rep("red", 32), "yellow"),
                  edge.col="darkgray")

sna::gplot.target(g, betweenness(g), main="Betweenness",
                  circ.lab = FALSE, circ.col="skyblue",
                  usearrows = FALSE,
                  vertex.col=c("blue", rep("red", 32), "yellow"),
                  edge.col="darkgray")

sna::gplot.target(g, evcent(g), main="Eigenvector",
                  circ.lab = FALSE, circ.col="skyblue",
                  usearrows = FALSE,
                  vertex.col=c("blue", rep("red", 32), "yellow"),
                  edge.col="darkgray")
```

# Example: Karate Club of Zachary

  ----------------------------------------------------
  Network Statistic         Zachary's Karate Club      
  ------------------------- -------------------------- 
  Avg. Degree               4.588                      
   
  Avg. Closeness Cen.       0.005                     
    
  Avg. Betweenness Cen.     26.194                     
    
  Avg. Eigenvector Cen.     0.377                      
    
  ------------------------- --------------------------

# Graph Models 

 - A *graph model* takes in fixed parameters and generates a graph that vary in structure with each iteration
 
 - Equivalently, it is a collection, or *ensemble* of graphs, denoted by

 $$\{\mathbb{P}_{\theta}(G), G \in \mathcal{G}: \theta \in \Theta\}$$
 - $\mathcal{G}$ is a collection or ensemble of possible graphs, $P_\theta$ is a *probability distribution* on the random graph $G$, and $\theta$ is a vector of parameters that describe the graphs that $G$ can be, ranging over possible parameters in $\Theta$

# Erdős–Rényi Model (1959)

 - Model with parameters: $N_V$, and $N_{E}$ or $p$

 - Model of the form $G(N_{V}, p)$ and $G(N_{V}, N_{E})$

```{r, fig.align = 'center', eval=TRUE, include = FALSE}
set.seed(499)

g1.er <- erdos.renyi.game(n = 10, p = 0.25)
g2.er <- erdos.renyi.game(n = 10, p = 0.25)

par(mfrow=c(1,2))
plot(g1.er, vertex.size=20, vertex.label.cex = 0.75)
plot(g2.er, vertex.size=20, vertex.label.cex = 0.75)
```

# Properties of the Erdős–Rényi Model

 - Short average path lenths
 
 - Low clustering coefficient 
 
 - For the $G(N_V, p)$ model, a particular simple graph $g$ with exactly $N_V$ vertices has probability $P(G = g) = p^{N_E}(1 - p)^{\left({N_V \choose 2} - N_E \right)}$

# Watts-Strogatz Model (1998)

 - Model with parameters: $N_V$, $r$, $p$
 
```{r, echo = FALSE}
set.seed(499)

g1.ws <- watts.strogatz.game(dim = 1, size = 10, nei = 2, p = 0)

g2.ws <- watts.strogatz.game(dim = 1, size = 10, nei = 2, p = 0)
# the effect of adding one edge to the circular arrangement
randomedgepairs <- sample(1:10, 8, replace=TRUE)
g2.ws <- simplify(add_edges(g2.ws, randomedgepairs))

par(mfrow=c(1,2))
plot(g1.ws, vertex.size = 20, vertex.label.cex = 0.75)
plot(g2.ws, vertex.size = 20, vertex.label.cex = 0.75, layout=layout_in_circle)

```
 
# Properties of the Watts-Strogatz Model

 - High clustering coefficient 
 
 - Small average path length 

# Exponential random graph models (ERGMs) I

 - Exponential random graph models (ERGMs) are a class of models that can be used to generate probability distributions
 
 - Flexible in design; we can decide our parameters 
 
 - Conduct goodness-of-fit tests for model assessment
 
# Exponential random graph models (ERGMs) II

 - The general form for an ERGM is as follows: 
 
 $$P_{\theta, \mathcal{G}}(\textbf{G} = \textbf{g}) = \frac {exp(\theta^{T}\textbf{s(g)})} {\kappa(\theta, \mathcal{G})},\textbf{g} \in \mathcal{G}$$
 
 - $\textbf{Y}$ is the random variable representing a random graph and $\textbf{g}$ is the particular adjacency matrix we observe. $\textbf{s(g)}$ is the vector of model statistics for $\textbf{g}$, $\theta$ is the vector of coefficients for those statistics, and $\kappa(\theta, \mathcal{G})$ is the quantity in the numerator summed over all possible networks

# Properties of ERGMs I

 - Deriving the Erdős-Rényi Model from ERGMs
 
 - Suppose we have a particular graph $g$ and the only statistic we have is $L(G)$, the number of edges in $g$
 
 $$P_{\theta, \mathcal{G}}(g) = \frac {exp(\theta_{L}L(g))} {\sum_{g' \in \mathcal{G}}^{} exp(\theta_{L}L(g'))} = \frac {exp(\theta_{L}L(g))} {\kappa(\theta, \mathcal{G})}, g \in \mathcal{G}$$
 
# Properties of ERGMs II

 - Consider the probability distribution for a particular graph $g$ with $N_E$ edges again (from the Erdős-Rényi model). Using the fact that $N_{E} = L(g)$, taking the equation as a power of base $e$, we get the following: 
 
$$
\begin{aligned}
P(g) &= p^{N_E}(1 - p)^{\left({N_V \choose 2} - N_E \right)} \\
&= p^{L(g)}(1 - p)^{\left(\frac {N_{V}(N_{V} - 1)} {2} - L(g) \right)} \\
&= \left( \frac {p} {1-p} \right)^{L(g)}(1 - p)^{\frac {N_{V}(N_{V} - 1)} {2}} \\
&= exp \left(L(g)log \left(\frac {p} {1-p} \right) - \frac {N_{V}(N_{V} - 1)} {2} log \left( \frac {p} {1-p} \right) \right) \\
&= exp(\theta_{L}L(g) - c) \\
&= \frac {exp(\theta_{L}L(g))} {exp(c)},
\end{aligned} 
$$

# Data Set

 - A component of the Facebook network
 
 - $4039$ vertices and $88234$ edges 
 
 - Simple, connected, and undirected 

# Generating Random Graphs

 - Create graphs of parable magnitude using certain parameters 
 
 - Simulate $1000$ random graphs for and record network statatistics for each graph
 
 - Create distribution of these values and/or see a table of averages
 
 - Compare network statistics of the Facebook network with that of the graphs we generated

# The Picture in Mind

```{r, echo=FALSE}
set.seed(490)

numsim <- 100 

g.er.karmod.coef <- rep(NA, numsim)
g.er.karmod.apl <- rep(NA, numsim)
g.er.karmod.dia <- rep(NA, numsim)

for (i in 1:numsim) {
  # n = number of vertices, p = probability of a link
  #p is caluclated as number of edges over number of possible edges
  #p is then (34 choose 2) since there is the possibility of a link between
  g.er.karmod <- erdos.renyi.game(n = vcount(karate), p = (ecount(karate)/choose(vcount(karate), 2)))
  g.er.karmod.coef[i] <- transitivity(g.er.karmod)
  g.er.karmod.apl[i] <- average.path.length(g.er.karmod)
  g.er.karmod.dia[i] <- diameter(g.er.karmod)
}
```

```{r, echo=FALSE, fig.align='center', fig.height=2, fig.width=5, warning=FALSE}
gplot1 <- ggplot(data = as.data.frame(g.er.karmod.coef), 
       aes(x = g.er.karmod.coef)) +
  geom_histogram(binwidth=0.0065, 
       colour="blue", 
       fill="lavender") + 
  geom_vline(xintercept = transitivity(karate), colour="red") + 
  labs(title="Simulated 
       Clustering Coefficients") + 
  xlab("Clustering Coefficient") + 
  ylab("Frequency") + 
  theme(legend.position="none", 
        plot.title = element_text(size=6),
        axis.text = element_text(size=4),
        axis.title = element_text(size=4)) 


gplot2 <- ggplot(data = as.data.frame(g.er.karmod.apl), aes(x = g.er.karmod.apl)) +
  geom_histogram(binwidth=0.031,
                 colour="blue",
                 fill = "lavender") + 
  geom_vline(xintercept = average.path.length(karate), colour="red") + 
  labs(title="Simulated 
       Average Path Lengths") + 
  xlab("Average Path Length") +
  ylab("") + 
  theme(legend.position="none", 
        plot.title = element_text(size=6),
        axis.text = element_text(size=4),
        axis.title = element_text(size=4)) 

gplot3 <- ggplot(data = as.data.frame(g.er.karmod.dia), aes(x = g.er.karmod.dia)) +
  geom_histogram(binwidth=0.12,
                 colour="blue",
                 fill = "lavender") + 
  geom_vline(xintercept = diameter(karate), colour="red") + 
  labs(title="Simulated Diameters") + 
  xlab("Diameter") + 
  ylab("") + 
  theme(legend.position="none", 
        plot.title = element_text(size=6),
        axis.text = element_text(size=4),
        axis.title = element_text(size=4))

grid.arrange(gplot1, gplot2, gplot3, ncol=3)
```

# Erdős–Rényi Model

 - $N_{V}$ and $p$
 - The number of vertices is $4039$
 - Estimate the probability by taking the number of observed edges and dividing by the number of possible edges $\hat{p} = \frac {88234} {{4039 \choose 2}} = 0.011$ 
 - For every possible edge among the $4039$ vertices, determine if an edge will form based on the estimated probability

# Watts-Strogatz Model 

 - $N_{V}$, $r$, $p$
 - Will not use $p$
 - Start with a lattice with $4039$ vertices
 - Randomly add $88234 - 4039 = 84195$ edges until we have approximately the same number as our observed network. 
 - Assign a number of edges to the vertices equal to the smallest degree observed in our Facebook network; $1$
 - Simplify our simulated graph to eliminate multi-edges and loop

# Results for Erdős–Rényi and Watts-Strogatz Models

  ----------------------------------------------------------------------------------------------
  Network Statistic     Observed     Erdős-Rényi                   Watts-Strogatz
  --------------------- ------------ ----------------------------- -----------------------------
  Transitivity           0.617        0.0108 $\pm$ 0.0001           0.0107 $\pm$ 9.135e-05
    
  Average Path Length    4.338        2.606 $\pm$ 0.002             2.6093 $\pm$ 0.0002           
    
  Diameter               17           3.96 $\pm$ 0.21               3.95 $\pm$ 0.22
    
  Avg. Degree Cen.       43.691       43.69 $\pm$ 0.14              43.45 $\pm$ 0.01 
    
  Avg. Betweenness Cen.  2072.642     3242 $\pm$ 4                  3249.2 $\pm$ 0.4 
    
  Avg. Closeness Cen.    8.881e-08    9.507e-05 $\pm$ 7.230e-08     9.494e-05 $\pm$ 7.319e-09 
    
  Avg. Eigenvector Cen.  0.040        0.620 $\pm$ 0.022             0.6235 $\pm$ 0.0227       
  ---------------------- ------------ ----------------------------- -----------------------------

# ERGMs

 - Four different ERGMs--labeled as ERGM 1a, ERGM 2a, ERGM 2b, and ERGM 3a
 
 - ERGM 1a: one parameter: edges 
 
 - ERGM 2a: two parameters: edges and triangles
 
 - ERGM 2b: two parameters: edges and k-stars (of size 3)
 
 - ERGM 3a: three parameters: edges, triangles, and k-stars (of size 3)
 
 - For each random graph, calculate the networks statistics of interest 

# Results for ERGMs I

  -------------------------------------------------------------------------------------------
  Network Statistic      Observed    ERGM1a                     ERGM2a
  ---------------------  ----------- -------------------------- -----------------------------
  Transitivity           0.617       0.3696 $\pm$ 0.0012        0.4823 $\pm$ 0.0020
    
  Average Path Length    4.338       2.885 $\pm$ 0.004          3.052 $\pm$ 0.0086         
    
  Diameter               17          5.216 $\pm$ 0.412          6.098 $\pm$ 0.035
    
  Avg. Degree            43.691      43.66 $\pm$ 0.052          44.54 $\pm$ 0.03 
    
  Avg. Betweenness Cen.  2072.642    3805 $\pm$ 9               4140 $\pm$ 18 
    
  Avg. Closeness Cen.    8.881e-08   8.286e-05 $\pm$ 8.353e-06  6.008e-05 $\pm$ 1.514e-05 
    
  Avg. Eigenvector Cen.  0.040       0.0417 $\pm$ 0.0008        0.0410 $\pm$ 3.577e-05      
  ---------------------- ----------- -------------------------- -----------------------------

# Results for ERGMs II

  -------------------------------------------------------------------------------------------
  Network Statistic      Observed    ERGM2b                     ERGM3a
  ---------------------- ----------- -------------------------- -----------------------------
  Transitivity           0.617       0.3787 $\pm$ 0.0013        0.4891 $\pm$ 0.0020
    
  Average Path Length    4.338       2.851 $\pm$ 0.003          3.0562 $\pm$ 0.0079  
    
  Diameter               17          5.095 $\pm$ 0.293          6.161 $\pm$ 0.3677
    
  Avg. Degree Cen.       43.691      44.06 $\pm$ 0.057          44.52 $\pm$ 0.034
    
  Avg. Betweenness Cen.  2072.642    3736 $\pm$ 6               4148 $\pm$ 16
    
  Avg. Closeness Cen.    8.881e-08   8.558e-05 $\pm$ 6.025e-06  5.875e-05 $\pm$ 1.5035e-05
    
  Avg. Eigenvector Cen.  0.040       0.0392 $\pm$ 0.0002        0.0410 $\pm$ 3.451e-05      
  ---------------------- ----------- -------------------------- -----------------------------

# Conclusions and Future Work 

 - The models were bad, so now what? 
 
 - Choose other models
 
 - Choose different network statistics 
 
 - Choose other data sets

# Implications

 - Understand the flow of information 

 - Better access to jobs through networking 
 
 - Better leads to resources in research 
 
 - Improving traffic 
 
 - Understanding biological systems

# Sources I

Baumer, B., Kaplan, D., & Horton, N. (2017). Modern Data Science With R: With Digital Download. Taylor & Francis. Retrieved from https://books.google.com/books?id=Gv1nvgAACAAJ

Butts, C. T., & others. (2008). Social network analysis with sna. Journal of Statistical Software, 24 (6), 1–51.

Butts, C., Hunter, D., Handcock, M. S., Morris, M., Krivtisky, P. N., Almquist, Z.,. . . Bender de-Moll, S. (2015, June). Introduction to Exponential-family Random Graph (ERG or p*) modeling with ergm. Retrieved from https://statnet.org/trac/raw-attachment/wiki/Sunbelt2015/ergm_tutorial.pdf

Hunter, D. R., Handcock, M. S., Butts, C. T., Goodreau, S. M., Morris, & Martina. (2008). Ergm: A package to fit, simulate and diagnose exponential-family models for networks. Journal of Statistical Software, 24 (3), 1–29.

# Sources II

Leskovec, J., Chakrabarti, D., Kleinberg, J., & Faloutsos, C. (2005). Realistic, Mathematically Tractable Graph Generation and Evolution, Using Kronecker Multiplication. In Knowledge Discovery in Databases: PKDD 2005 (pp. 133–145). Springer, Berlin, Heidelberg. Retrieved from http://link.springer.com/chapter/10.1007/11564126_17

Jackson, M. O. (2013). Social and economic networks: Models and analysis. StanfordUniversity; Coursera Course Lecture.

Kolaczyk, E. D. (2009). Statistical Analysis of Network Data: Methods and Models. Springer Science & Business Media.

Kolaczyk, E. D., & Csárdi, G. (2014).Statistical Analysis of Network Data with R.Springer.

Lazega, E., & Pattison, P. E. (1999). Multiplexity, generalized exchange and cooperation in organizations: A case study. Social Networks, 21 (1), 67–90.

# Sources III

Leskovec, J., & Krevl, A. (2014, June). SNAP Datasets: Stanford large network
dataset collection. http://snap.stanford.edu/data.

Leskovec, J., Chakrabarti, D., Kleinberg, J., & Faloutsos, C. (2005). Realistic, Mathematically Tractable Graph Generation and Evolution, Using Kronecker Multiplication. In Knowledge Discovery in Databases: PKDD 2005 (pp. 133–145). Springer, Berlin, Heidelberg. Retrieved from http://link.springer.com/chapter/10.1007/11564126_17

Leskovec, J., Chakrabarti, D., Kleinberg, J., Faloutsos, C., & Ghahramani, Z. (2010).

Kronecker Graphs: An Approach to Modeling Networks. Journal of Machine Learning Research, 11 (Feb), 985–1042. Retrieved from http://www.jmlr.org/55papers/v11/leskovec10a.html

# Sources IV

Newman, M. (2010). Networks: An Introduction. OUP Oxford

Travers, J., & Milgram, S. (1967). The small world problem. Phychology Today, 1,
61–67.

Watts, D. J., & Strogatz, S. H. (1998). Collective dynamics of “small-world” networks. Nature, 393 (6684), 440–442. http://doi.org/10.1038/30918

Zachary, W. W. (1977). An information flow model for conflict and fission in small groups. Journal of Anthropological Research, 33 (4), 452–473. http://doi.org/10.1086/jar.33.4.3629752
 