---
title: "A Simulation Study Using Random Graph Models to Fit Social Networks"
author: |
 | Levi Lee 
 | Advisor: Amy Wagaman
 | Amherst College
date: April 12, 2017

output:
  beamer_presentation:
    theme: "Warsaw"
    colortheme: "orchid"
    fonttheme: "professionalfonts"
    fig_caption: true
---

```{r, include=FALSE}
library(sand)
library(igraph)
library(network)
library(sna)
library(statnet)
library(ergm)
library(xtable)
library(mdsr)
library(gridExtra)

options(xtable.comment = FALSE)
```

# A Highly Connected World 

 - Networks are everywhere 
 
 - Informational, biological, technological, and social applications 
 
 - Focus is on social networks

# History 

 - Travers & Milgram (1967)
  
 - Letter correspondence between strangers in Nebraska and Massachusetts 
  
 - Required only six people on average for the letter to be delivered 
  
 - "Six degrees of separation"

# Social Networks

 - Involve a lot of people 

 - Form highly condensed groups

 - Have relatively short distances between people 

# Question

 - Can we simulate this? If so, how? 
 
 - Conduct a simulation study utilizing different graph models 

# Basic Terminlogy

 - A *graph*, denoted $G$ or $G(V, E)$, consists of a set of *vertices* $i, j, k, ... \in V$ and a set of *edges* $\{i, j\}, \{i, k\}, \{j, k\}... \in E$. 
 
 - Our focus is on *simple*, *undirected*, and *connected* graphs. 

# Other Representations of Graphs

 - *Adjacency matrix*: denoted $\textbf{A}$ is an $N_V$ x $N_V$ matrix where each element denotes the existence of edges between pairs

$$
A_{ij} = \begin{cases}
    1 & \text{if } \{i, j\} \in E \\
    0 & \text{otherwise.} 
  \end{cases}
$$

 - *Edge list*: two-column list of all the edges in a graph denoted by their corresponding vertices

# Example: Karate Club of Zachary (1977) 

```{r, echo=FALSE, fig.height=4.5, fig.align='center'}
data(karate)
plot(karate)
```

# Example: Lazega's Law Firm (1999) 

```{r lawyerplot, message = FALSE, eval = TRUE,  echo = FALSE, fig.align = 'center'}
# Office location indicated by color.
colbar <- c("red", "dodgerblue", "goldenrod")
v.colors <- colbar[V(lazega)$Office]
# Type of practice indicated by vertex shape.
v.shapes <- c("circle", "square")[V(lazega)$Practice]
# Vertex size proportional to years with firm.
v.size <- 3.5*sqrt(V(lazega)$Years)
# Label vertices according to seniority.
v.label <- V(lazega)$Seniority
# Reproducible layout.
set.seed(42)
l <- layout.fruchterman.reingold(lazega)
plot(lazega, layout=l, vertex.color=v.colors,
     vertex.shape=v.shapes, vertex.size=v.size,
     vertex.label=v.label)
```

# Network Statistics 

 - Analogous to statistics seen in elementary statistics

 - Characterize a given network with a numerical summaries

# Transitivity/Clustering Coefficient
 
 - Ratio of triangles to connected triples
   
 - *Triangle*: three vertices connected by three edges 
 
 - *Connected triple*: three vertices connected by two edges

 $$C = \frac {(\text{number of triangles}) \times 3} {\text{number of connected triples}}$$
 
 - Local versus global

# Notions of Distance 

*Average path length*: average of the shortest paths of all distinct pairs of vertices in the network

*Diameter*: longest of all the shortest paths between distinct pairs of vertices

# Example: Karate Club and Law Firm 
  ------------------------------------------------------------------
  Network Statistic       Zachary's Karate Club  Lazega's Law Firm 
  ---------------------- ---------------------- ------------------
  Transitivity (Global)  0.256                   0.389
  
  Transitivity (Local)   0.588                   0.487
    
  Avg. Path Length       2.408                   2.144
  
  Diameter               13                      5
  ---------------------- ---------------------- ------------------
  
# Centrality

 - Measure of importance for each vertex in the graph
 
 - Many different types of centralities exist 

# Degree Centrality 

 - Based on the number of edges that are connected to a vertex 

 - Most common form of centrality

# Closeness Centrality 

 - Measures how close a vertex is to other vertices based on the inverse of the total distance of the vertex from all others
 
 $$c_{Cl}(i) = \frac {1} {\sum_{j \in V}^{} d(i, j)}$$
 
 - $dist(i,j)$ is the geodesic distance between the vertices $i,j \in V$

# Betweenness Centrality 

 - Measures the extent to which a vertex is located between other pairs of vertices
 
 $$c_{B}(i) = \sum_{g \neq h \neq i \in V}^{} \frac {\sigma(g,h|i)} {\sigma(g,h)}$$

 - $\sigma(g,h|i)$ is the total number of shortest paths between $g$ and $h$ that pass through $i$, and $\sigma(g,h) = \sum_{i \in V}^{} \sigma(g,h|i)$

# Eigenvector Centrality 

 - Based on the idea of "status," "prestige," or "rank;" the more central the neighbors of a vertex are, the more central that vertex itself is
 
 $$c_{Ei}(i) = \alpha \sum_{\{i,j\} \in E}^{} c_{Ei} (u)$$

 - $c_{Ei} = (c_{Ei}(1), ..., c_{Ei}(N_V))^T$ is the solution to the eigenvalue problem $\textbf{A}\textbf{c}_{Ei} = \alpha^{-1}\textbf{c}_{Ei}$, where $\textbf{A}$ is the adjacency matrix for network graph $G$.

# Example: Karate Club

```{r, echo = FALSE, fig.align = 'center', warning = FALSE}
A <- get.adjacency(karate, sparse=FALSE)
library(network)
g <- network::as.network.matrix(A)
library(sna)
par(mfrow=c(2,2))
sna::gplot.target(g, degree(g), main="Degree",
                  circ.lab = FALSE, circ.col="skyblue",
                  usearrows = FALSE,
                  vertex.col=c("blue", rep("red", 32), "yellow"),
                  edge.col="darkgray")


sna::gplot.target(g, closeness(g), main="Closeness",
                  circ.lab = FALSE, circ.col="skyblue",
                  usearrows = FALSE,
                  vertex.col=c("blue", rep("red", 32), "yellow"),
                  edge.col="darkgray")

sna::gplot.target(g, betweenness(g), main="Betweenness",
                  circ.lab = FALSE, circ.col="skyblue",
                  usearrows = FALSE,
                  vertex.col=c("blue", rep("red", 32), "yellow"),
                  edge.col="darkgray")

sna::gplot.target(g, evcent(g), main="Eigenvector",
                  circ.lab = FALSE, circ.col="skyblue",
                  usearrows = FALSE,
                  vertex.col=c("blue", rep("red", 32), "yellow"),
                  edge.col="darkgray")
```

# Example: Karate Club and Law Firm

  --------------------------------------------------------------
  Network Statistic    Zachary's Karate Club  Lazega's Law Firm   
  -------------------- ---------------------- ------------------
  Degree               4.588                  6.389    
   
  Closeness Cen.       0.005                  0.007   
    
  Betweenness Cen.     26.194                 17.833
    
  Eigenvector Cen.     0.377                  0.458
  -------------------- ---------------------- -------------------

# Graph Models 

 - A *graph model* takes in fixed parameters and generates a graph that varies in structure with each iteration
 
 - Equivalently, it is a collection, or *ensemble* of graphs, denoted by

 $$\{\mathbb{P}_{\theta}(G), G \in \mathcal{G}: \theta \in \Theta\}$$
  - $\mathcal{G}$ is a collection or ensemble of possible graphs, $P_\theta$ is a *probability distribution* on the random graph $G$, and $\theta$ is a vector of parameters that describe the graphs that $G$ can be, ranging over possible parameters in $\Theta$

# Erdős–Rényi Model (1959)

 - Model with parameters: $N_V$, and $N_{E}$ or $p$
 - Model of the form $G(N_{V}, p)$ and $G(N_{V}, N_{E})$

```{r, fig.align = 'center', eval=TRUE, include = FALSE}
set.seed(499)

g1.er <- erdos.renyi.game(n = 10, p = 0.25)
g2.er <- erdos.renyi.game(n = 10, p = 0.25)

par(mfrow=c(1,2))
plot(g1.er, vertex.size=20, vertex.label.cex = 0.75)
plot(g2.er, vertex.size=20, vertex.label.cex = 0.75)
```

# Properties of the Erdős–Rényi Model

 - Short average path lengths
 - Low clustering coefficient 
 - For the $G(N_V, p)$ model, a particular simple graph $g$ with exactly $N_V$ vertices has probability $P(G = g) = p^{N_E}(1 - p)^{\left({N_V \choose 2} - N_E \right)}$

# Watts-Strogatz Model (1998)

 - Model with parameters: $N_V$, $r$, $p$
 
```{r, echo = FALSE}
set.seed(499)

g1.ws <- watts.strogatz.game(dim = 1, size = 10, nei = 2, p = 0)

g2.ws <- watts.strogatz.game(dim = 1, size = 10, nei = 2, p = 0)
# the effect of adding one edge to the circular arrangement
randomedgepairs <- sample(1:10, 8, replace=TRUE)
g2.ws <- simplify(add_edges(g2.ws, randomedgepairs))

par(mfrow=c(1,2))
plot(g1.ws, vertex.size = 20, vertex.label.cex = 0.75)
plot(g2.ws, vertex.size = 20, vertex.label.cex = 0.75, layout=layout_in_circle)

```
 
# Properties of the Watts-Strogatz Model

 - High clustering coefficient 
 - Small average path length 

# Exponential random graph models (ERGMs) I

 - Exponential random graph models (ERGMs) are a class of models that can be used to generate probability distributions
 - Flexible in design; we can decide our parameters 
 - Conduct goodness-of-fit tests for model assessment
 
# Exponential random graph models (ERGMs) II

 - The general form for an ERGM is as follows: 
 $$P_{\theta, \mathcal{G}}(\textbf{G} = \textbf{g}) = \frac {exp(\theta^{T}\textbf{s(g)})} {\kappa(\theta, \mathcal{G})},\textbf{g} \in \mathcal{G}$$
 
 - $\textbf{G}$ is the random variable representing a random graph and $\textbf{g}$ is the particular adjacency matrix we observe. $\textbf{s(g)}$ is the vector of model statistics for $\textbf{g}$, $\theta$ is the vector of coefficients for those statistics, and $\kappa(\theta, \mathcal{G})$ is the quantity in the numerator summed over all possible networks

# Properties of ERGMs I

 - Deriving the Erdős-Rényi Model from ERGMs
 
 - Suppose we have a particular graph $g$ and the only statistic we have is $L(G)$, the number of edges in $g$
 
 $$P_{\theta, \mathcal{G}}(g) = \frac {exp(\theta_{L}L(g))} {\sum_{g' \in \mathcal{G}}^{} exp(\theta_{L}L(g'))} = \frac {exp(\theta_{L}L(g))} {\kappa(\theta, \mathcal{G})}, g \in \mathcal{G}$$
 
# Properties of ERGMs II

 - Consider the probability distribution for a particular graph $g$ with $N_E$ edges again (from the Erdős-Rényi model). Using the fact that $N_{E} = L(g)$, taking the equation as a power of base $e$, we get the following: 
 
$$
\begin{aligned}
P(g) &= p^{N_E}(1 - p)^{\left({N_V \choose 2} - N_E \right)} \\
&= p^{L(g)}(1 - p)^{\left(\frac {N_{V}(N_{V} - 1)} {2} - L(g) \right)} \\
&= \left( \frac {p} {1-p} \right)^{L(g)}(1 - p)^{\frac {N_{V}(N_{V} - 1)} {2}} \\
&= exp \left(L(g)log \left(\frac {p} {1-p} \right) - \frac {N_{V}(N_{V} - 1)} {2} log \left(\frac {1} {1 - p} \right) \right) \\
\end{aligned} 
$$

# Properties of ERGMs III

$$
\begin{aligned}
P(g) &= exp(\theta_{L}L(g) - c) \\
&= \frac {exp(\theta_{L}L(g))} {exp(c)}
\end{aligned} 
$$

 - This is the form of the ERGM seen earlier 
 
  $$P_{\theta, \mathcal{G}}(g) = \frac {exp(\theta_{L}L(g))} {\kappa(\theta, \mathcal{G})}, g \in \mathcal{G}$$
 
 - Constant $c$ is exactly the normalizing constant we need to scale this into a probability distribution

# Data Set

 - A component of the Facebook network
 - $4039$ vertices and $88234$ edges 
 - Simple, connected, and undirected 
 
# The Facebook Component 

![Overview of our component of the Facebook network](figure/31fbelplot.pdf)

# Generating Random Graphs

 - Create graphs of parable magnitude using certain parameters 
 - Simulate $1000$ random graphs for and record network statistics for each graph
 - Create distribution of these values and/or see a table of averages
 - Compare network statistics of the Facebook network with that of the graphs we generated

# The Picture in Mind

```{r, echo=FALSE}
set.seed(490)

numsim <- 100 

g.er.karmod.coef <- rep(NA, numsim)
g.er.karmod.apl <- rep(NA, numsim)
g.er.karmod.dia <- rep(NA, numsim)

for (i in 1:numsim) {
  # n = number of vertices, p = probability of a link
  #p is caluclated as number of edges over number of possible edges
  #p is then (34 choose 2) since there is the possibility of a link between
  g.er.karmod <- erdos.renyi.game(n = vcount(karate), p = (ecount(karate)/choose(vcount(karate), 2)))
  g.er.karmod.coef[i] <- transitivity(g.er.karmod)
  g.er.karmod.apl[i] <- average.path.length(g.er.karmod)
  g.er.karmod.dia[i] <- diameter(g.er.karmod)
}
```

```{r, echo=FALSE, fig.align='center', fig.height=2, fig.width=5, warning=FALSE}
gplot1 <- ggplot(data = as.data.frame(g.er.karmod.coef), 
       aes(x = g.er.karmod.coef)) +
  geom_histogram(binwidth=0.0065, 
       colour="blue", 
       fill="lavender") + 
  geom_vline(xintercept = transitivity(karate), colour="red") + 
  labs(title="Simulated 
       Clustering Coefficients") + 
  xlab("Clustering Coefficient") + 
  ylab("Frequency") + 
  theme(legend.position="none", 
        plot.title = element_text(size=6),
        axis.text = element_text(size=4),
        axis.title = element_text(size=4)) 


gplot2 <- ggplot(data = as.data.frame(g.er.karmod.apl), aes(x = g.er.karmod.apl)) +
  geom_histogram(binwidth=0.031,
                 colour="blue",
                 fill = "lavender") + 
  geom_vline(xintercept = average.path.length(karate), colour="red") + 
  labs(title="Simulated 
       Average Path Lengths") + 
  xlab("Average Path Length") +
  ylab("") + 
  theme(legend.position="none", 
        plot.title = element_text(size=6),
        axis.text = element_text(size=4),
        axis.title = element_text(size=4)) 

gplot3 <- ggplot(data = as.data.frame(g.er.karmod.dia), aes(x = g.er.karmod.dia)) +
  geom_histogram(binwidth=0.12,
                 colour="blue",
                 fill = "lavender") + 
  geom_vline(xintercept = diameter(karate), colour="red") + 
  labs(title="Simulated Diameters") + 
  xlab("Diameter") + 
  ylab("") + 
  theme(legend.position="none", 
        plot.title = element_text(size=6),
        axis.text = element_text(size=4),
        axis.title = element_text(size=4))

grid.arrange(gplot1, gplot2, gplot3, ncol=3)
```

# Erdős–Rényi Model

 - $N_{V}$ and $p$
 - Number of vertices is $N_{V} = 4039$
 - Estimate the probability by taking the number of observed edges and dividing by the number of possible edges $\hat{p} = \frac {88234} {{4039 \choose 2}} = 0.011$ 
 - For every possible edge among the $4039$ vertices, determine if an edge will form based on the estimated probability

# Watts-Strogatz Model 

 - $N_{V}$, $r$, $p$
 - Will not use $p$
 - Start with a circular model with $4039$ vertices
 - Assign a number of edges to the vertices equal to the smallest degree observed in our Facebook network; $deg_{min} = 1$
 - Randomly add $88234 - 4039 = 84195$ edges until we have the same number as our observed network
 - Simplify our simulated graph to eliminate multi-edges and loops

# Results for Erdős–Rényi and Watts-Strogatz Models

  ----------------------------------------------------------------------------------------------
  Network Statistic     Observed     Erdős-Rényi                   Watts-Strogatz
  ------------------ ------------ ----------------------------- -----------------------------
  Transitivity        0.617        0.0108 $\pm$ 0.0001           0.0107 $\pm$ 9e-05
    
  Avg. Path Length    4.338        2.606 $\pm$ 0.002             2.609 $\pm$ 0.0002           
    
  Diameter            17           3.96 $\pm$ 0.21               3.95 $\pm$ 0.22
    
  Degree Cen.         43.691       43.69 $\pm$ 0.14              43.45 $\pm$ 0.01 
    
  Betweenness Cen.    2072.642     3242 $\pm$ 4                  3249.2 $\pm$ 0.4 
    
  Closeness Cen.      8.881e-08    9.507e-05 $\pm$ 7e-08         9.494e-05 $\pm$ 7e-09 
    
  Eigenvector Cen.    0.040        0.620 $\pm$ 0.022             0.6235 $\pm$ 0.0227       
  ------------------- ------------ ----------------------------- -----------------------------

# ERGMs

 - Four different ERGMs--labeled as ERGM 1a, ERGM 2a, ERGM 2b, and ERGM 3a
 - ERGM 1a: one parameter: edges 
 - ERGM 2a: two parameters: edges and triangles
 - ERGM 2b: two parameters: edges and k-stars (of size 3)
 - ERGM 3a: three parameters: edges, triangles, and k-stars (of size 3)
 - Estimate the coefficients ($\theta$'s) of the probability model 
 - Simulate

# Results for ERGMs I

  ----------------------------------------------------------------------------------------
  Network Statistic   Observed    ERGM1a                     ERGM2a
  ------------------  ----------- -------------------------- -----------------------------
  Transitivity        0.617       0.3696 $\pm$ 0.0012        0.4823 $\pm$ 0.0020
    
  Avg. Path Length    4.338       2.885 $\pm$ 0.004          3.052 $\pm$ 0.009         
    
  Diameter            17          5.216 $\pm$ 0.412          6.098 $\pm$ 0.035
    
  Degree              43.691      43.66 $\pm$ 0.05           44.54 $\pm$ 0.03 
    
  Betweenness Cen.    2072.642    3805 $\pm$ 9               4140 $\pm$ 18 
    
  Closeness Cen.      8.881e-08   8.286e-05 $\pm$ 8.35e-06   6.008e-05 $\pm$ 1.514e-05 
    
  Eigenvector Cen.    0.040       0.0417 $\pm$ 0.0008        0.0410 $\pm$ 5e-05      
  ------------------- ----------- -------------------------- ----------------------------

# Results for ERGMs II

  --------------------------------------------------------------------------------------
  Network Statistic      Observed    ERGM2b                     ERGM3a
  ------------------  ----------- -------------------------- -----------------------------
  Transitivity        0.617       0.3787 $\pm$ 0.0013        0.4891 $\pm$ 0.0020
    
  Avg. Path Length    4.338       2.851 $\pm$ 0.003          3.056 $\pm$ 0.008 
    
  Diameter            17          5.095 $\pm$ 0.293          6.161 $\pm$ 0.368
    
  Degree Cen.         43.691      44.06 $\pm$ 0.06           44.52 $\pm$ 0.03
    
  Betweenness Cen.    2072.642    3736 $\pm$ 6               4148 $\pm$ 16
    
  Closeness Cen.      8.881e-08   8.558e-05 $\pm$ 6.03e-06   5.875e-05 $\pm$ 1.5035e-05
 
  Eigenvector Cen.    0.040       0.0392 $\pm$ 0.0002        0.0410 $\pm$ 3e-05      
  ------------------- ----------- -------------------------- -----------------------------

# Conclusions and Future Work 

 - The models were bad, so now what? 
 - Choose other models
 - Choose different network statistics 
 - Choose other data sets

# Implications

 - Understand the flow of information
 - Better access to jobs through networking 
 - Better leads to resources in research 
 - Improving traffic 
 - Understanding biological systems
 - Product recommendations

# Sources I
 - Baumer, B., Kaplan, D., & Horton, N. (2017). Modern Data Science With R: With Digital Download. Taylor & Francis. 
 - Butts, C. & others. (2008). Social network analysis with sna. 
 - Butts, C. & others. (2015, June). Introduction to Exponential-family Random Graph modeling with ergm.
 - Hunter, D. & others. (2008). Ergm: A package to fit, simulate and diagnose exponential-family models for networks. 
 - Leskovec, J. & others.(2005). Realistic, Mathematically Tractable Graph Generation and Evolution, Using Kronecker Multiplication.
 - Jackson, M. O. (2013). Social and economic networks: Models and analysis.
 - Kolaczyk, E. D. (2009). Statistical Analysis of Network Data: Methods and Models.
 - Kolaczyk, E. D. & Csárdi, G. (2014). Statistical Analysis of Network Data with R.

# Sources II
 - Lazega, E. & Pattison, P. E. (1999). Multiplexity, generalized exchange and cooperation in organizations: A case study.
 - Leskovec, J. & Krevl, A. (2014, June). SNAP Datasets: Stanford large network dataset collection. 
 - Leskovec, J. & others. (2005). Realistic, Mathematically Tractable Graph Generation and Evolution, Using Kronecker Multiplication.
 - Leskovec, J. & others. (2010). Kronecker Graphs: An Approach to Modeling Networks.
 - Newman, M. (2010). Networks: An Introduction. 
 - Travers, J., & Milgram, S. (1967). The small world problem.
 - Watts, D. J., & Strogatz, S. H. (1998). Collective dynamics of “small-world” networks. 
 - Zachary, W. W. (1977). An information flow model for conflict and fission in small groups.