---
title: "A Short Introduction to Networks and Model Comparisons"
author: |
 | Levi Lee 
 | Advisor: Amy Wagaman
 | Amherst College
date: April 12, 2017

output:
  beamer_presentation:
    theme: "Szeged"
    colortheme: "orchid"
    fonttheme: "professionalfonts"
---

```{r, include=FALSE}
library(sand)
library(igraph)
library(network)
library(sna)
library(statnet)
library(ergm)
library(xtable)
library(mdsr)
library(gridExtra)

options(xtable.comment = FALSE)
```

# Background

# A Highly Connected World 

 - Networks are everywhere 

 - Technological, biological, social, informational 

# History 

Travers & Milgram (1967)
  
 - Letter correspondence between strangers in Nebraska and Massachusetts 
  
 - Overall, it took only around six people for the letter to be delivered 
  
 - "Six degrees of separation"

# Social Networks

 - Involves a lot of people 

 - Highly condensed groups

 - Relatively short distances between people 

# Question

 - Can we simulate this? 

# Basic Definitons 

 - A *graph*, denoted $G(V, E)$, consists of a set of *vertices* $i, j, k, ... \in V$ and a set of *edges* $\{i, j\}, \{j, k\}... \in E$. 
 
 - *Order*: number of vertices in $G$, denoted $|V| = N_V $ 
 
 - *Size* : number of edges in $G$, denoted $|E| = N_E$.

 - Our focus is on *simple*, *undirected*, and *connected* graphs. 

# Other Representations of Graphs

 - *Adjacency matrix*: denoted $\textbf{A}$ is an $N_V$ x $N_V$ matrix each element denotes the exitence of edges between pairs of vertices such that 

\[ A_{ij} = \begin{cases}
    1 & \text{if } \{i, j\} \in E \\
    0 & \text{otherwise.} 
  \end{cases}
\]

 - *Edge list*: two-column list of all the edges in a graph denoted by their corresponding vertices present

# Example: Karate Club of Zachary (1977) 

```{r, echo=FALSE, fig.height=5, fig.align='center'}
data(karate)
plot(karate)
```

# Example: Lazega's Law Firm (1999)

```{r lawyerplot, message = FALSE, eval = FALSE, fig.align='center', echo = FALSE}
# Office location indicated by color.
colbar <- c("red", "dodgerblue", "goldenrod")
v.colors <- colbar[V(lazega)$Office]
# Type of practice indicated by vertex shape.
v.shapes <- c("circle", "square")[V(lazega)$Practice]
# Vertex size proportional to years with firm.
v.size <- 3.5*sqrt(V(lazega)$Years)
# Label vertices according to seniority.
v.label <- V(lazega)$Seniority
# Reproducible layout.
set.seed(42)
l <- layout.fruchterman.reingold(lazega)
plot(lazega, layout=l, vertex.color=v.colors,
     vertex.shape=v.shapes, vertex.size=v.size,
     vertex.label=v.label)
```

# Network Statistics 

 - Analogous to statisitcs seen in elementary statistics

 - Characterizes a given network

# Transitivity/Clustering Coefficient

 - Probability that adjacent vertices of a particular vertex are connected
 
 - Ratio of connected triples to all possible triangles in the graph
   
 - *Triangle* (or *triad*): three vertices connected by three edges 
 
 - *Connected triple*: three vertices connected by two edges

$$C = \frac {(\text{number of triangles}) \times 3} {\text{number of connected triples}$$

# Distances 

*Average path length*: average of the shortest paths of all distinct pairs of vertices in the network

*Diameter*: longest of all the shortest paths between distinct pairs of vertices

# Example: Karate Club of Zachary (1977) 

[TABLE OF VALUES]

# Example: Lazega's Law Firm

[TABLE OF VALUES]

# Centrality

 - Measure of importance for each vertex in the graph
 
 - Many different types of centralities exist 

# Degree Centrality 

 - Vertices with higher vertex degrees are considered to be more central to the network than those with lower vertex degrees

# Betweenness Centrality 

 - 

# Closeness Centrality 

 - 

# Eigenvector Centrality 

# Example: Karate Club of Zachary (1977) 

# Graph Models 

# Erdős–Rényi Model

# Watts-Strogatz Model 

# ERGMs

# Simulation Study 

# Data Set

# Erdős–Rényi Model

# Watts-Strogatz Model 

# ERGMs

# Conclusion

# Results 

# Implications

# Sources

